% Initialisation
\documentclass[a4paper]{article}

% Enforce UTF8 encoding
\usepackage[utf8]{inputenc}

% Header and footer
\usepackage{fancyhdr}
 
\pagestyle{fancy}
\fancyhead[L]{}
\fancyhead[R]{SENG4800 - Individual project proposal \\ \today}
\fancyfoot[C]{\thepage}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

% Document title and author
\title{3D visualisation of large data in a multi-platform, touch-based environment}
\author{Monica Olejniczak}

% Paragraphing
\setlength{\parindent}{0em}
\setlength{\parskip}{1em}

% Todo notes

\usepackage[disable]{todonotes}
\presetkeys{todonotes}{inline}{}

% References
\usepackage{natbib}
\renewcommand{\bibname}{References}

\begin{document}

	% Title page
	\makeatletter
	\vspace*{\fill}
	\begin{center}
		\Huge\textbf{\textsf{\@title}}
		\rule{\textwidth}{1pt}
		\LARGE\textsc{Individual project} \\
		\Large\textsc{University of Newcastle}
		\rule{\textwidth}{1pt}
		\Large\textbf{\textsf{\@author}}
	\end{center}
	\vspace*{\fill}
	\makeatother
	\newpage
	
	\section{Project description} {
	
		Data visualisation is the presentation of data in a visual context and is particularly important in data analytics since humans are able to recognise patterns, trends and correlations more easily \citep{grinstein2002introduction}.
		
		This project aims to develop touch-based visualisations in three-dimensions and will focus on using large data in a generic way to demonstrate general applicability. The visualisations will provide touch-capability so they are able to be viewed on a smart device.
	
	}
	
	\section{Basic deliverables} {

		\begin{itemize}
			\item In-depth topic literature review.
			\item Build the implementation for a single visualisation with a small dataset.
			\item Basic interaction with 3D content - rotate, pan and zoom.
			\item Port the visualisation to a touch-interface.
		\end{itemize}
	
	}
	
	\section{Intermediate deliverables} {
	
		\begin{itemize}
			\item Incorporate additional interactions, including data filtering and volume selection.
			\item Build the implementation for another visualisation with a large dataset.
			\item Apply with one or more large datasets.
		\end{itemize}	
	
	}
	
	\section{Advanced deliverables} {
	
		\begin{itemize}
			\item Incorporate touch gestures.
			\item Analyse the computational power needed for the visualisations.
			\item Perform a user study to compare what types of visualisations are suited for particular purposes and its impact on user interactions.
			\item Benchmark the scalability of the visualisations over datasets.
			\item Integrate with data from the Final Year Project.
		\end{itemize}
	
	}
	
	\section{Literature review} {
	
		\subsection{Issues with big data} {
		
			\begin{itemize}
				\item \citet{kaisler2013big}
				\item \citet{katal2013big}		
			\end{itemize}		
		
		}
		
		\subsection{Information visualisation} {
		
			\begin{itemize}
				\item \citet{robertson1993information}
				\item \citet{rohrer1997web}
				\item \citet{ware2012information}
				\item \citet{duval2011attention}
			\end{itemize}
		
		}
		
		\subsection{Visualising big data} {
		
		}
		
	}
	
	\section{Approval} {
	
		\vspace{2em}\par
		\textbf{Student:} 
		\vspace{3em}\par
		\textbf{Supervisor:}
		
	}
	
	\newpage
	\bibliography{references}
	\bibliographystyle{apalike}
	
\end{document}
